project-plan

### 2025-09-26 project-plan.qmd
Revision based on feedback

Among trauma-related conditions, traumatic brain injury (TBI) is a major cause of death and disability worldwide and
constitutes a significant public health burden.(3) TBI is the result of an external force causing neurological
impairments that may be either transient or permanent.(4) Furthermore, TBI should not be regarded as a single event
but rather as a dynamic process. The initial trauma **often triggers a secondary pathophysiological cascade**, including
ischemia, oedema, and metabolic disturbances, that develop over time and critically influence long-term outcomes. (5)
As such, patients **with TBI** often require admission to the intensive care unit (ICU).

@@

(At Karolinska University Hospital, the setting for this study, similar principles are seen in the Traumahandboken.
For severe TBI, defined as GCS < 8, early airway protection with intubation in unconscious patients, rapid CT
imaging, timely administration of tranexamic acid, and early neurosurgical consultation are recommended. In cases
where there are signs of impending herniation, the handbook advises measures such as controlled hyperventilation and
osmotherapy with mannitol or hypertonic saline. All this is always in conjunction with urgent neurosurgical
evaluation. (Källa Karolinska handboken))  <!-- Remove paragraph? MGW: Yes, I think so.-->

@@

<!--Previous papers at Karolinska University Hospital have contributed to this field by describing OFIs across different
aspects of trauma care. In 2023, Hussein A. et al. analysed patient- and process-related factors during trauma
resuscitation and concluded that variables such as time to first CT and injury severity were strongly associated with
OFI (17). Cao ES. et al. continued the focus on ICU cohorts, presenting that 9.9% ICU-admitted adult trauma patients
experienced at least one OFI (1). Most recently, Szolnoky K et al. assessed OFI patterns in trauma cohorts and
demonstrated that, although overall incidence declined over time, patients with TBI remained likely to have
preventable shortcomings in care (18).-->

@@

Study Design
-----------
A single-centre, registry-based retrospective cohort study was conducted at Karolinska University Hospital in Solna
(NKS). The trauma registry for NKS is a comprehensive trauma registry that **is part of** the Swedish National Trauma
Registry, SweTrau, and is compliant with the Utstein template for uniform reporting after trauma (Källa  Utstein
Template). As such, the registry prospectively records patient demographics, prehospital and emergency department
physiology, times of key interventions such as imaging, surgery and transfer, injury severity scores, and in-hospital outcome.

@@

Participants
------------
All adult patients with TBI recorded in the Karolinska trauma registry between 2012 and
2023 who also had been screened for opportunity for improvements **were included in this study**. Furthermore, TBI was defined as (a GCS score of ≤8
combined with an Abbreviated Injury Scale ≥3 for the head region ….  ) Patients were eligible irrespective of level <!-- Definition found in Elins paper, same applies?--><!-- Yes -->
of care.

@@

Logistic regression was used to examine the association between patient and process-related factors and OFI.
Bivariable logistic regression to assess unadjusted associations between patient and process factors and OFI was
initially performed. Variables of clinical relevance and those significant in bivariable analyses were then entered
into a multivariable logistic regression model to estimate adjusted associations. <!-- MGW: I think you should decide which ones to include in the multivariable model from the start, regardless of their significance in the bivariable analysis --> Results are reported as Odds Ratio
(OR) with 95% confidence interval (CI), and a p-value of <0.05 was used as the threshold for statistical
significance. To ensure objectivity, all statistical analyses were initially performed on synthetic data before being
applied to the combined trauma registry and trauma care quality database. Statistical analyses were conducted in R
version x. <!-- I assume we will do both bivariable and multivariable regression? -->

>>>>

Among trauma-related conditions, traumatic brain injury (TBI) is a major cause of death and disability worldwide and
constitutes a significant public health burden (3). TBI is the result of an external force causing neurological
impairments that may be either transient or permanent (4). Furthermore, TBI should not be regarded as a single event
but rather as a dynamic process. The initial trauma often triggers a secondary pathophysiological cascade, including ischemia, oedema, and metabolic disturbances, that develop over time and critically influence long-term outcomes (5). As such, patients with TBI often require admission to the intensive care unit (ICU).

@@

@@

@@

Study Design
-----------
A single-centre, registry-based retrospective cohort study was conducted at Karolinska University Hospital in Solna
(NKS). The trauma registry for NKS is a comprehensive trauma registry that is part of the Swedish National Trauma
Registry, SweTrau, and is compliant with the Utstein template for uniform reporting after trauma (Källa  Utstein
Template). As such, the registry prospectively records patient demographics, prehospital and emergency department
physiology, times of key interventions such as imaging, surgery and transfer, injury severity scores, and in-hospital outcome.

@@

Participants
------------
All adult patients with TBI recorded in the Karolinska trauma registry between 2012 and 2023 who also had been
screened for opportunity for improvements were included in this study. Furthermore, TBI was defined as a GCS score of
≤8 combined with an Abbreviated Injury Scale ≥3 for the head region. Patients were eligible irrespective of level of
care.

@@

Logistic regression was used to examine the association between patient and process-related factors and OFI.
Bivariable logistic regression to assess unadjusted associations between patient and process factors and OFI was
initially performed. Variables of clinical relevance and those significant in bivariable analyses were then entered
into a multivariable logistic regression model to estimate adjusted associations. <!-- MGW: I think you should decide which ones to include in the multivariable model from the start, regardless of their significance in the bivariable analysis --> Results are reported as Odds Ratio
(OR) with 95% confidence interval (CI), and a p-value of <0.05 was used as the threshold for statistical
significance. To ensure objectivity, all statistical analyses were initially performed on synthetic data before being
applied to the combined trauma registry and trauma care quality database. Statistical analyses were conducted in R
version x.
